---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# some.models

<!-- badges: start --> [![Build Status](https://travis-ci.org/<github username>/<repo name>.svg?branch=master)](https://travis-ci.org/<github username>/<repo name>)
[![Codecov test coverage](https://codecov.io/gh/benbren/some.models/branch/master/graph/badge.svg)](https://codecov.io/gh/benbren/some.models?branch=master)
<!-- badges: end -->

The goal of some.models is to allow the user to use and understand both linear regression and logistic regression in a user friendly way. In this package, we create functions with output that is limited to metrics that are important when beginning to learn about these regression techniques. Compared to lm and glm,. these functions are easy to understand and interpret, and allow for easy, real-world, prediction capabilites. 

## Installation

You can install the released version of some.models from GitHub. First, you should run 
``` r 
.rs.restartR()

```

followed by 

``` r
install.packages('devtools')
devtools::install_github('benbren/some.models')
library("some.models")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(some.models)

data(swiss)
X = swiss$Catholic
y = swiss$Infant.Mortality

fit = linear(X,y)

names(fit) # you can now choose what to do with all these things! 

ucla =  read.csv("https://stats.idre.ucla.edu/stat/data/binary.csv")

X = cbind(ucla$gpa)

y = ucla$admit

fit_logistic = logistic(X,y)

names(fit_logistic) # more things to choose from


```

These are basic examples, please use 

```r
?linear
```

to see more complex examples and how to use these functions in a more complete way. 
